---
title: "ningze_sun_hw1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First of all: loading all library needed for this homework readr, stringr, hash, and define all the functions.

The reason of using hash function is that I would like a hash-like function for counting based on my coding style.

```{r library, warning=0}
library('readr')
library('stringr')
library('hash')
split_inside = function(i){
  t = str_split(str_replace_all(i, '[[:punct:]]', ""), ' ')[[1]]
  return(t[t != ""])
}


hash_count = function(split){
  count = hash()
  for (i in split){
    if (is.null(count[[i]])){
      count[[i]] = 1
    }
    else{
      count[[i]] = count[[i]] + 1
    }
  }
  return(count)
}

path = getwd()
if (path != 'E:/class/statistic/BIO384K/ningze-sun-ada-homework-1'){
  setwd('E:/class/statistic/BIO384K/ningze-sun-ada-homework-1')
}
```


## Challenge 1:

Use read_dim() to load data file “darwin.txt” into a character vector, where each element is a paragraph.

```{r read_file}
file = read_delim('darwin.txt', '\n')[[1]]
```

Calculate and print out how many paragraphs are there in the chapter and print out the 34th paragraph

```{r part_1}
print(paste("In darwin.txt, there are", toString(length(file)), "paragraphs"))
print(paste("The 34th paragraph is: ", file[[34]]))
```

Split into new vector, split comprising 1-word character strings using split_inside function with str_split and str_replace_all to remove all unword-character such as . , 

These functions are defined at very beginning for saving coding writing.

```{r part_2}
split = c()
for (i in file){
  split = c(split, split_inside(i))
}

```

print how many unique words in this vector by unique function; Count each word appearing using hash_count function by hash and loop. 

Then loop the hash by its keys to count most common word, appear 1 time, 5 or more times.

```{r part_3}
print(paste("There are", toString(length(unique(split))), "unique words in this vector"))
count = hash_count(split)
max_name = ""
max_value = 0
equal = 0
larger = 0
for (i in keys(count)){
  if (count[[i]] == 1){
    equal = equal + 1
  }
  if (count[[i]] >= 5){
    larger = larger + 1
  }
  if (count[[i]] > max_value){
    max_value = count[[i]]
    max_name = i
  }
}
print(paste("The most common word are", max_name))
print(paste(toString(equal), "words are appear 1 time"))
print(paste(toString(larger), "words are appear more than 5 time"))

```

Extract the entire text of the last paragraph to a new variable, final_quote;

Split this vector into a new vector, split_quote, comprising 1-word vectors by split_inside and use seq to select every 3 third vector and use sort to rearrange the elements.

```{r part_4}
final_quote = file[[length(file)]]
split_quote = split_inside(final_quote)
every_third = split_quote[seq(3, length(split_quote), 3)]
print("Every third elemnt from this vector are: ")
print(every_third)
print("The reverse alphabetical order of every_third are ")
print(sort(every_third, decreasing = TRUE))

```


## Challenge 2:


Create the temperature and city using c(), and use names to associate data with city and use [] and : to print required temperature

```{r part_6}
t = c(35, 88, 42, 84, 81, 30)
city= c('Beijing', 'Lagos', 'Paris', 'Rio de Janeiro', 'San Juan', 'Toronto')
names(t) = city
print("The first 3 cities in the list are")
print(t[city[1:3]])
print("The temperatures of Paris and San Juan are")
print(t[city[c(3, 5)]])

```

## Challenge 3:

Create a 8 row by 20 column matrix,consisting descending order from 159 to 0 filling matrix column-wise.

```{r part_7}
m1 = matrix(seq(159,0), nrow = 8, ncol = 20, byrow= TRUE)
print(paste("extract and print the element in row 5, column 2:", toString(m1[5,2])))
print("extract and print the contents of rows 5 to 7 of the matrix.")
print(m1[5:7,])
m2 = m1[3:4,6:9]
print(paste("The class of variable is", toString(class(m2))))
print(paste("The mode of this variable is", toString(mode(m2))))

```


## Challenge 4:

Construct a 4-dimensional, 400-element array (5 x 5 x 4 x 4), consisting of even integers from 2 to 800 

```{r part_8}
a = array(seq(800, 1, -2), dim = c(5,5,4,4))
print(a[1, 1, 1, 2])
print(a[2, 3, 2, ])
print(a[1:5, 1:5, 3, 3])

```







## Challenge 5:

Create a list representing the primate taxonomy outlined. Use hard coding for each family, superfamily et.

Use double bracket notation to extract the New World monkeys (Parvorder Platyrrhini) into a new variable, print the class and mode

Use $ notation to extract the tarisers

```{r part_9}
# superfamily
Lorisoidea = c('Lorisidae', 'Galagidae')
Lemuroidea = c('Cheirogaleidae', 'Lepilemuridae', 'Indriidae', 'Lemuridae', 'Daubentoniidae')
Tarsioidea = c('Tarsiidae')
Ceboidea = c("Cebidae", "Atelidae", 'Pitheciidae')
Hominoidea = c('Hylobatidae', 'Hominidae')
Cercopithecoidea = c('Cercopithecidae')
# infraorder
Lorisiformes = list(Lorisoidea=Lorisoidea)
Lemuriformes = list(Lemuroidea=Lemuroidea)
Tarsiiformes = list(Tarsioidea=Tarsioidea)
# Parvorder
Platyrrhini = list(Ceboidea=Ceboidea)
Catarrhini = list(Hominoidea=Hominoidea, Cercopithecoidea=Cercopithecoidea)
Simiiformes = list(Platyrrhini=Platyrrhini, Catarrhini=Catarrhini)
# suborder
Strepsirhini = list(Lorisiformes=Lorisiformes, Lemuriformes=Lemuriformes)
Haplorhini = list(Tarsiiformes=Tarsiiformes,Simiiformes=Simiiformes)
# order
Primates = list(Strepsirhini=Strepsirhini,Haplorhini=Haplorhini)
platyrrhines = Primates[['Haplorhini']][['Simiiformes']][['Platyrrhini']]
print(paste("The class of variable is", toString(class(platyrrhines))))
print(paste("The mode of this variable is", toString(mode(platyrrhines))))
tariser = Primates$Haplorhini$Tarsiiformes$Tarsioidea
print(tariser)

```

